Interval Probability and Measurement for Fuzzy Quantum Theories

### 2016-10-10 prob.tex
prob.tex \subsection{Classical Interval-valued probability measures} Update the type of the operators...

As motivated above, an event $E_{1}$ may have an interval of probability
$[l_{1},r_{1}]$. Assume that another disjoint event $E_{2}$ has
interval probability $[l_{2},r_{2}]$, what is the interval probability
of the event $E_{1}\cup E_{2}$? The answer is somewhat subtle: although
it is possible to use the sum of the intervals $[l_{1}+l_{2},r_{1}+r_{2}]$
as the combined probability, one can do find a much tighter interval
if information \emph{against} the event (i.e., information about the
complement event) is also taken into consideration. Formally, for
a general event $E$ with probability $[l,r]$, the evidence that
contradicts $E$ is an evidence supporting the complement of $E$.
The complement of $E$ must therefore have probability $\left[1-r,1-l\right]$
which we abbreviate $\left[1,1\right]\tensor*[_{\mathscr{I}}]{-}{}\left[l,r\right]$,
where the preposing $\mathscr{I}$ specifies we subtract intervals.
Given a collection of intervals $\mathscr{I}$, an $\mathscr{I}$\textendash interval-valued
probability measure is a function $\bar{\mu}:\events\rightarrow\mathscr{I}$
such that~\cite{JamisonLodwick2004}:\footnote{The right-end of the interval-valued probability measure is a special
case of (sub-additive) capacity~\cite{Choquet1954,Graf1980,Goodman1997}
or outer measure~\cite{Folland1999}. The left-end of the interval-valued
probability measure is a special case of (super-additive cooperative)
game~\cite{Shapley1971,Grabisch2016}.} 
\begin{itemize}
\item $\bar{\mu}(\emptyset)=[0,0]$. 
\item $\bar{\mu}(\Omega)=[1,1]$. 
\item For every collection of pairwise disjoint events $\left\{ E_{i}\right\} _{i=0}^{N-1},\left\{ E_{i}'\right\} _{i=0}^{N'-1}\subseteq\events$
with $\Omega=\left(\bigcup_{i=0}^{N-1}E_{i}\right)\bigcup\left(\bigcup_{i=0}^{N'-1}E_{i}'\right)$,
we have 
\[
\bar{\mu}\left(\bigcup_{i=0}^{N-1}E_{i}\right)\subseteq\left[\max\left\{ 1-\sideset{_{\mathbb{R}}}{_{i=0}^{N'-1}}\sum r_{i}',\sideset{_{\mathbb{R}}}{_{i=0}^{N-1}}\sum l_{i}\right\} ,\min\left\{ 1-\sideset{_{\mathbb{R}}}{_{i=0}^{N'-1}}\sum l_{i}',\sideset{_{\mathbb{R}}}{_{i=0}^{N-1}}\sum r_{i}\right\} \right]\textrm{ ,}
\]
where $\bar{\mu}\left(E_{i}\right)=[l_{i},r_{i}]$ and $\bar{\mu}\left(E_{i}'\right)=[l_{i}',r_{i}']$
for all $i$.
\end{itemize}
In order to understand this definition, we tear the last condition
apart and provide an equivalent definition of interval-valued probability
measures.

@@

\begin{example}[Two-coin experiment with interval probability]
\label{ex3} We split the unit interval $[0,1]$ in the following
four closed sub-intervals: $[0,0]$ which we call \imposs, $[0,\frac{1}{2}]$
which we call \unlikely, $[\frac{1}{2},1]$ which we call \likely,
and $[1,1]$ which we call \necess. Using these new values, we can
modify the probability measure of Ex.~\ref{ex1} by mapping each
numeric value to the smallest sub-interval containing it to get the
following: 
\[
\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\bar{\mu}(\emptyset) & = & \imposs\\
\bar{\mu}(\{HH\}) & = & \unlikely\\
\bar{\mu}(\{HT\}) & = & \imposs\\
\bar{\mu}(\{TH\}) & = & \likely\\
\bar{\mu}(\{TT\}) & = & \imposs\\
\bar{\mu}(\{HH,HT\}) & = & \unlikely\\
\bar{\mu}(\{HH,TH\}) & = & \necess\\
\bar{\mu}(\{HH,TT\}) & = & \unlikely
\end{array} & \begin{array}{rcl}
\bar{\mu}(\{HT,TH\}) & = & \likely\\
\bar{\mu}(\{HT,TT\}) & = & \imposs\\
\bar{\mu}(\{TH,TT\}) & = & \likely\\
\bar{\mu}(\{HH,HT,TH\}) & = & \necess\\
\bar{\mu}(\{HH,HT,TT\}) & = & \unlikely\\
\bar{\mu}(\{HH,TH,TT\}) & = & \necess\\
\bar{\mu}(\{HT,TH,TT\}) & = & \likely\\
\bar{\mu}(\{HH,HT,TH,TT\}) & = & \necess
\end{array}\end{array}
\]
Despite the absence of any numeric information, the probability measure
is quite informative: it reveals that the second coin is double-headed
and that the first coin is biased. To understand the $\subseteq$-condition,
consider the following calculation: 
\begin{eqnarray*}
 &  & \bar{\mu}(\{HH\})+\bar{\mu}(\{HT\})+\bar{\mu}(\{TH\})+\bar{\mu}(\{TT\})\\
 & = & \imposs+\unlikely+\imposs+\likely\\
 & = & \left[0,0\right]+\left[0,\frac{1}{2}\right]+\left[0,0\right]+\left[\frac{1}{2},1\right]=\left[\frac{1}{2},\frac{3}{2}\right]
\end{eqnarray*}
If we were to equate $\bar{\mu}(\Omega)$ with the sum of the individual
probabilities, we would get that $\bar{\mu}(\Omega)=\left[\frac{1}{2},\frac{3}{2}\right]$.
However, using the fact that $\bar{\mu}(\emptyset)=\imposs$, we have
$\bar{\mu}\left(\Omega\right)=1-\bar{\mu}\left(\emptyset\right)=\necess=[1,1]$.
This interval is tighter and a better estimate for the probability
of the event $\Omega$, and of course it is contained in $[\frac{1}{2},\frac{3}{2}]$.
However it is only possible to exploit the information about the complement
when all four events are combined. Thus the $\subseteq$-condition
allows us to get an estimate for the combined event from each of its
constituents and then gather more evidence knowing the aggregate event.\qed\end{example}

>>>>

As motivated above, an event $E_{1}$ may have an interval of probability
$[l_{1},r_{1}]$. Assume that another disjoint event $E_{2}$ has
interval probability $[l_{2},r_{2}]$, what is the interval probability
of the event $E_{1}\cup E_{2}$? The answer is somewhat subtle: although
it is possible to use the sum of the intervals $[l_{1}\tensor*[_{\mathbb{R}}]{+}{}l_{2},r_{1}\tensor*[_{\mathbb{R}}]{+}{}r_{2}]$
as the combined probability, one can do find a much tighter interval
if information \emph{against} the event (i.e., information about the
complement event) is also taken into consideration. Formally, for
a general event $E$ with probability $[l,r]$, the evidence that
contradicts $E$ is an evidence supporting the complement of $E$.
The complement of $E$ must therefore have probability $\left[1\tensor*[_{\mathbb{R}}]{-}{}r,1\tensor*[_{\mathbb{R}}]{-}{}l\right]$
which we abbreviate $\left[1,1\right]\tensor*[_{\mathscr{I}}]{-}{}\left[l,r\right]$,
where the preposing $\mathscr{I}$ specifies we subtract intervals.
Given a collection of intervals $\mathscr{I}$, an $\mathscr{I}$\textendash interval-valued
probability measure is a function $\bar{\mu}:\events\rightarrow\mathscr{I}$
such that~\cite{JamisonLodwick2004}:\footnote{The right-end of the interval-valued probability measure is a special
case of (sub-additive) capacity~\cite{Choquet1954,Graf1980,Goodman1997}
or outer measure~\cite{Folland1999}. The left-end of the interval-valued
probability measure is a special case of (super-additive cooperative)
game~\cite{Shapley1971,Grabisch2016}.} 
\begin{itemize}
\item $\bar{\mu}(\emptyset)=[0,0]$. 
\item $\bar{\mu}(\Omega)=[1,1]$. 
\item For every collection of pairwise disjoint events $\left\{ E_{i}\right\} _{i=0}^{N-1},\left\{ E_{i}'\right\} _{i=0}^{N'-1}\subseteq\events$
with $\Omega=\left(\bigcup_{i=0}^{N-1}E_{i}\right)\bigcup\left(\bigcup_{i=0}^{N'-1}E_{i}'\right)$,
we have 
\[
\bar{\mu}\left(\bigcup_{i=0}^{N-1}E_{i}\right)\subseteq\left[\max\left\{ 1-\sideset{_{\mathbb{R}}}{_{i=0}^{N'-1}}\sum r_{i}',\sideset{_{\mathbb{R}}}{_{i=0}^{N-1}}\sum l_{i}\right\} ,\min\left\{ 1-\sideset{_{\mathbb{R}}}{_{i=0}^{N'-1}}\sum l_{i}',\sideset{_{\mathbb{R}}}{_{i=0}^{N-1}}\sum r_{i}\right\} \right]\textrm{ ,}
\]
where $\bar{\mu}\left(E_{i}\right)=[l_{i},r_{i}]$ and $\bar{\mu}\left(E_{i}'\right)=[l_{i}',r_{i}']$
for all $i$.
\end{itemize}
In order to understand this definition, we tear the last condition
apart and provide an equivalent definition of interval-valued probability
measures.

@@

\begin{example}[Two-coin experiment with interval probability]
\label{ex3} We split the unit interval $[0,1]$ in the following
four closed sub-intervals: $[0,0]$ which we call \imposs, $[0,\frac{1}{2}]$
which we call \unlikely, $[\frac{1}{2},1]$ which we call \likely,
and $[1,1]$ which we call \necess. Using these new values, we can
modify the probability measure of Ex.~\ref{ex1} by mapping each
numeric value to the smallest sub-interval containing it to get the
following: 
\[
\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\bar{\mu}(\emptyset) & = & \imposs\\
\bar{\mu}(\{HH\}) & = & \unlikely\\
\bar{\mu}(\{HT\}) & = & \imposs\\
\bar{\mu}(\{TH\}) & = & \likely\\
\bar{\mu}(\{TT\}) & = & \imposs\\
\bar{\mu}(\{HH,HT\}) & = & \unlikely\\
\bar{\mu}(\{HH,TH\}) & = & \necess\\
\bar{\mu}(\{HH,TT\}) & = & \unlikely
\end{array} & \begin{array}{rcl}
\bar{\mu}(\{HT,TH\}) & = & \likely\\
\bar{\mu}(\{HT,TT\}) & = & \imposs\\
\bar{\mu}(\{TH,TT\}) & = & \likely\\
\bar{\mu}(\{HH,HT,TH\}) & = & \necess\\
\bar{\mu}(\{HH,HT,TT\}) & = & \unlikely\\
\bar{\mu}(\{HH,TH,TT\}) & = & \necess\\
\bar{\mu}(\{HT,TH,TT\}) & = & \likely\\
\bar{\mu}(\{HH,HT,TH,TT\}) & = & \necess
\end{array}\end{array}
\]
Despite the absence of any numeric information, the probability measure
is quite informative: it reveals that the second coin is double-headed
and that the first coin is biased. To understand the $\subseteq$-condition,
consider the following calculation: 
\begin{eqnarray*}
 &  & \bar{\mu}(\{HH\})\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}(\{HT\})\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}(\{TH\})\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}(\{TT\})\\
 & = & \imposs\tensor*[_{\mathscr{I}}]{+}{}\unlikely\tensor*[_{\mathscr{I}}]{+}{}\imposs\tensor*[_{\mathscr{I}}]{+}{}\likely\\
 & = & \left[0,0\right]\tensor*[_{\mathscr{I}}]{+}{}\left[0,\frac{1}{2}\right]\tensor*[_{\mathscr{I}}]{+}{}\left[0,0\right]\tensor*[_{\mathscr{I}}]{+}{}\left[\frac{1}{2},1\right]=\left[\frac{1}{2},\frac{3}{2}\right]
\end{eqnarray*}
If we were to equate $\bar{\mu}(\Omega)$ with the sum of the individual
probabilities, we would get that $\bar{\mu}(\Omega)=\left[\frac{1}{2},\frac{3}{2}\right]$.
However, using the fact that $\bar{\mu}(\emptyset)=\imposs$, we have
$\bar{\mu}\left(\Omega\right)=1\tensor*[_{\mathscr{I}}]{-}{}\bar{\mu}\left(\emptyset\right)=\necess=[1,1]$.
This interval is tighter and a better estimate for the probability
of the event $\Omega$, and of course it is contained in $[\frac{1}{2},\frac{3}{2}]$.
However it is only possible to exploit the information about the complement
when all four events are combined. Thus the $\subseteq$-condition
allows us to get an estimate for the combined event from each of its
constituents and then gather more evidence knowing the aggregate event.\qed\end{example}
